---
title: "Optimizations and parallel processing"
output: rmarkdown::html_vignette
description: |
  How can we evaluate candidate models in the shortest amount of time?
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Optimizations and parallel processing}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidymodels)
```

The tune package attempts to evaluate the candidate models in the shortest amount of time. This vignette is a summary of those approaches. 

## Sub-model speed-ups

For some types of models, such as boosted models or regularized models, the number of models that are actually _fit_ can be far less than the number of models evaluated. For example, suppose a boosted tree is fit with 1000 trees. Many boosting implementations let the user make predictions for any number of trees less than what was originally fit (1000 in this example). This "sub-model trick" can greatly speed up the training time for many models (e.g. see [this example](https://topepo.github.io/caret/using-your-own-model-in-train.html#illustrative-example-2-something-more-complicated---logitboost) in the `caret` documentation). 

In order to know what models allow this, the parsnip package contains a `multi_predict()` function that enables this feature. Printing the S3 methods for it lists the possible models:

```{r submodels}
library(tidymodels)
methods("multi_predict")

# There are arguments for the parameter(s) that can create multiple predictions.
# For xgboost, `trees` are cheap to evaluate: 
parsnip:::multi_predict._xgb.Booster %>% 
  formals() %>% 
  names()
```

The same feature does not exist for recipes though. 

## Expensive pre-processing 

When tuning a recipe and a model, it makes sense to avoid _recreating_ the recipe for each model. 

For example, suppose that [Isomap multi-dimensional scaling](https://en.wikipedia.org/wiki/Isomap) is used to pre-process the data prior to tuning a K-nearest neighbor regression:

```{r isomap, results="hide"}
data(Chicago)
iso_rec <- 
  recipe(ridership ~ ., data = Chicago) %>% 
  step_dummy(all_nominal()) %>% 
  step_isomap(all_predictors(), num_terms = tune())

knn_mod <- 
  nearest_neighbor(neighbors = tune(), weight_func = tune()) %>% 
  set_engine("kknn") %>% 
  set_mode("regression")
```

With the following grid:

```{r}
grid <- 
  parameters(num_terms(c(1, 9)), neighbors(), weight_func()) %>% 
  grid_regular(levels = c(5, 10, 7)) %>% 
  arrange(num_terms, neighbors, weight_func)
grid 
```

To evaluate these `r nrow(grid)` candidate models, we would have to compute the same recipe 70 times _per resample_. Since Isomap is expensive, this is really inefficient. 

`tune_grid()` determines when this occurs and fits all 70 candidate models for each unique configuration of the recipe. In essence, it nests the model parameters inside the unique parameters of the recipe:

```{r nest, echo = FALSE}
alt_grid <- tidyr::nest(grid, data = c(-num_terms))
alt_grid
```

When `parallel_over = "resamples"`, the default, only `r nrow(alt_grid)` recipes are prepared and, within each, all of the appropriate models are fit from the same recipe. In this example, once the recipe with `num_terms = 1` is created, the model parameters are iteratively tuned: 

```{r model-param, echo = FALSE}
alt_grid$data[[1]]
```

The same will be true for post-processing parameters being tuned. For each unique set of recipe and model parameters, the post-processing parameters will be evaluated without unnecessary re-fitting. 

Also, when using a model formula, the model matrix is only created once per resample. 

## Parallel processing

The tune package allows users, when possible, to use multiple cores or separate machines to fit models. The package is currently able to parallelize over either the resampling loop of grid search (via `parallel_over = "resamples"` in `control_grid()`, the default) or both the resampling and preprocessing loops (via `parallel_over = "everything"`). When `parallel_over = "everything"`, an outer parallel loop will iterate over resamples and an inner parallel loop will iterate over all unique combinations of preprocessor and model tuning parameters for that specific resample. This will result in the preprocessor being re-processed multiple times, but can be faster if that preprocessing is extremely fast.

tune supports parallel processing using the [future](https://www.futureverse.org/) framework. future supports a variety of technologies to share the computations and the choice of technology is determined by the chosen _plan_. To run tuning code in parallel, just provide a `plan()` and tune will take care of the rest. For example:

```{r future, eval = FALSE}
library(future)
plan(multisession)
```

### foreach (legacy)

Before the 1.2.0 release, tune supported parallelism using the foreach framework. Support has been deprecated for backends registered for foreach and will be fully removed in an upcoming release. foreach can use a variety of technologies to share the computations and the choice of technology is determined by which _parallel backend_ package is chosen (aka the `do{technology}` packages). For example, the doMC package uses the forking mechanism on Unix-like systems to split the computations across multiple cores. As of this writing, the backend packages are doMC, doMPI, doParallel, doRedis, doSNOW, and doAzureParallel (GitHub only).

Registering a parallel backend is also somewhat dependent of the package. For doParallel, one could use:

```{r doPar, eval = FALSE}
all_cores <- parallel::detectCores(logical = FALSE)

library(doParallel)
cl <- makePSOCKcluster(all_cores)
registerDoParallel(cl)
```

To **transition from foreach to future**, remove lines loading any foreach backend packages `library(do*)` as well as lines registering that technology with `registerDo*()`, and add the following lines:

```{r future2, eval = FALSE}
library(future)
plan(multisession)
```

Switch out `multisession` for another strategy if desired.

### In practice

One downside to parallel processing is that the different technologies handle inputs and outputs differently. For example, multicore forking _tends_ to carry the loaded packages and objects into the worker processes. Others do not. To make sure that the correct packages are loaded (but not attached) in the workers is to use the `pkg` option in `control_grid()`. 

Some helpful advice to avoid errors in parallel processing is to not use variables in the global environment. These may not be found when the code is run inside of a worker process. For example:

```{r par-tip, eval = FALSE}
num_pcs <- 3

recipe(mpg ~ ., data = mtcars) %>% 
  # Bad since num_pcs might not be found by a worker process
  step_pca(all_predictors(), num_comp = num_pcs)

recipe(mpg ~ ., data = mtcars) %>% 
  # Good since the value is injected into the object
  step_pca(all_predictors(), num_comp = !!num_pcs)
```

This issue is likely to occur if `dplyr::one_of()` is used as a sector.  


Also note that almost all of the logging provided by `tune_grid()` will not be seen when running in parallel. Again, this is dependent on the backend package and technology being used. 



