---
title: "Grid Search"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Grid Search}
output:
  knitr:::html_vignette:
    toc: yes
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidymodels)
library(tune)
library(mlbench)
library(kernlab)
library(crayon)
library(rlang)
```

To demonstrate model tuning, we'll use the Ionosphere data in the `mlbench` package:

```{r load-data}
library(mlbench)
data(Ionosphere)
```

There are 43 predictors and a factor outcome. Two of the predictors are factors (`V1` and `V2`) and the rest are numerics that have been scaled to a range of -1 to 1. Note that the two factor predictors have sparse distributions:

```{r factor-pred}
table(Ionosphere$V1)
table(Ionosphere$V2)
```

There's no point of putting `V2` into any model since is is a zero-variance predictor. `V1` is not but it _could_ be if the resampling process ends up sampling all of the same value. Is this an issue? It might be since the standard R formula infrastructure fails when there is only a single observed value:

```{r glm-fail, error=TRUE}
glm(Class ~ ., data = Ionosphere, family = binomial)

# Surprisingly, this doesn't help: 

glm(Class ~ . - V2, data = Ionosphere, family = binomial)
```

At a minimum, let's get rid of the most problematic variable:

```{r ion-rm}
Ionosphere <- Ionosphere %>% select(-V2)
```

## Inputs for the Search

To demonstrate, we'll fit a radial basis function support vector machine to these data and tune the SVM cost parameter and the $\sigma$ parameter in the kernel function:

```{r svm-mod}
svm_mod <-
  svm_rbf(cost = tune(), rbf_sigma = tune()) %>%
  set_mode("classification") %>%
  set_engine("kernlab")
```

In the code below, tuning will be demonstrated using a standard R formula as well as this recipe:

```{r rec}
iono_rec <-
  recipe(Class ~ ., data = Ionosphere)  %>%
  # In case V1 is has a single value sampled
  step_zv(all_predictors()) %>% 
  # convert it to a dummy variable
  step_dummy(V1) %>%
  # Scale it the same as the others
  step_range(matches("V1_"))
```

The only other required item for tuning is a resampling strategy as defined by an `rsample` object. Let's demonstrate using basic bootstrapping:

```{r rs}
set.seed(4943)
iono_rs <- bootstraps(Ionosphere, times = 30)
```

## Optional Inputs

An _optional_ step for model tuning is to specify which metrics should be computed using the out-of-sample predictions. For classification, the default is to calculate the log-likelihood statistic and overall accuracy. Instead of the defaults, the area under the ROC curve will be used. To do this, a `yardstick` function can be used to create a metric set:

```{r roc}
roc_vals <- metric_set(roc_auc)
```

If no grid or parameters are provided, a set of 10 are created using a space-filling design (via a Latin hypercube). A grid can be given in a data frame where the parameters are in columns and parameter combinations are in rows. Here, the default will be used.

Also, a control object can be passed that specifies different aspects of the search. Here, the verbose option is turned off. 

```{r ctrl}
ctrl <- grid_control(verbose = FALSE)
```

## Executing the Grid Using a Formula

First, the formula interface will be used:

```{r grid}
set.seed(35)
grid_form <- tune_grid(Class ~ ., model = svm_mod, rs = iono_rs, perf = roc_vals, control = ctrl)
grid_form
```

The `.metrics` column contains tibbles of the performance metrics for each tuning parameter combination:

```{r raw-metrics}
grid_form %>% select(.metrics) %>% slice(1) %>% pull(1)
```

To get the final resampling estimates, the `estimate` function can be used on the grid object:

```{r metric-estimates}
estimates <- estimate(grid_form)
estimates
```

The best combination was:

```{r sorted-metrics}
estimates %>% 
  arrange(desc(mean)) %>% 
  slice(1)
```

## Executing the Grid Using a Recipe

The same syntax is used but a recipe is passed in the first argument:


```{r recipe}
set.seed(325)
rec_form <- tune_grid(iono_rec, model = svm_mod, rs = iono_rs, perf = roc_vals, control = ctrl)
rec_form
```

The best setting here was:

```{r best-rec}
estimate(rec_form) %>% 
  arrange(desc(mean)) %>% 
  slice(1)
```

